TacticToe: Learning to Prove with Tactics

### 2018-02-20 tactictoe_jv.tex
minor
Minor grammar tweaks

\section{Introduction}
Many of the state-of-the-art interactive theorem provers (ITPs) such as
  \holfour~\cite{hol4}, \hollight~\cite{Harrison09hollight},
  \isabelle~\cite{isabelle}
  and \coq~\cite{coq-book} provide high-level parameterizable tactics for constructing proofs.
  Such tactics analyze the current proof state (goal and
  assumptions) and apply non-trivial proof transformations.
  Formalized proofs take advantage of different levels of automation: 
  specialized rules, theory-based strategies and general purpose startegies. 
  Thanks to the progress in proof 
  automation, developers can delegate more and more complicated proof 
  obligations to general purpose strategies. Those are implemented by automated 
  theorem provers(ATPs) such as \eprover~\cite{eprover}. The communications 
  between 
  an ITP and ATPs is made possible by a ``hammer''  
  system~\cite{hammers4qed,tgck-cpp15}. It acts as an interface by performing 
  premise selection, translation and proof reconstruction.
  Yet, ATPs are far from foolproof and more precise user-guidance, achieved 
  by applying a particular sequence of specialized rules, is almost always 
  necessary to develop a mathematical theory.

  In this work, we develop in the ITP \holfour a procedure to select suitable 
  tactics depending on the current proof state by learning 
  from previous proofs. Backed by this machine-learned guidance, our prover 
  \tactictoe relies on Monte 
  Carlo tree search~\cite{montecarlo} to find sequences of tactic applications 
  leading to an ITP proof.
  To this end, we \emph{extract and record} tactic invocations from the ITP
  proofs (Section~\ref{sec:recording}) and
  \emph{build efficient machine learning classifiers} based on such training
  examples (Section~\ref{s:prediction}).  The learned data serves as
  guidance for a \emph{Monte Carlo tree search} that explores different
  proof paths (Section~\ref{sec:proofsearch}). The result, if
  successful, is a verified human-level proof composed of \holfour
  tactics (Section~\ref{sec:proofdisplay}). The system is evaluated on a large 
  set of theorems originating
  from \holfour (Section~\ref{s:experiments}).

@@

We optimize \tactictoe by tuning parameters of six different techniques:
the timeout of tactics, orthogonalization, abstraction, MCTS policy, MCTS 
evaluation and ATP integration. During training, optional techniques such as 
orthogonalization may also be turned off completely.
\tactictoe also includes other important techniques which are run with their 
default parameters: feature extraction, prediction 
algorithms, number of premises for ATPs and MCTS evaluation radius (set by default to
10).
By choosing a set of parameters, we create a strategy for 
\tactictoe that is evaluated on our training set of 273 theorems. And the 
strategy with the highest number of proven theorems is selected for 
a full scale evaluation.

>>>>

\section{Introduction}
Many of the state-of-the-art interactive theorem provers (ITPs) such as
  \holfour~\cite{hol4}, \hollight~\cite{Harrison09hollight},
  \isabelle~\cite{isabelle}
  and \coq~\cite{coq-book} provide high-level parameterizable tactics for constructing proofs.
  Tactics analyze the current proof state (goal and
  assumptions) and apply non-trivial proof transformations.
  Formalized proofs take advantage of different levels of automation:
  specialized rules, theory-based strategies and general purpose startegies.
  Thanks to progress in proof
  automation, developers can delegate more and more complicated proof
  obligations to general purpose strategies. Those are implemented by automated
  theorem provers (ATPs) such as \eprover~\cite{eprover}. Communication
  between
  an ITP and ATPs is made possible by a ``hammer''
  system~\cite{hammers4qed,tgck-cpp15}. It acts as an interface by performing
  premise selection, translation and proof reconstruction.
  Yet, ATPs are far from foolproof and more precise user-guidance, achieved
  by applying a particular sequence of specialized rules, is almost always
  necessary to develop a mathematical theory.

  In this work, we develop in the ITP \holfour a procedure to select suitable
  tactics depending on the current proof state by learning
  from previous proofs. Backed by this machine-learned guidance, our prover
  \tactictoe relies on \emph{Monte
  Carlo tree search}~\cite{montecarlo} to find sequences of tactic applications
  leading to an ITP proof.
  To this end, we \emph{extract and record} tactic invocations from the ITP
  proofs (Section~\ref{sec:recording}) and
  \emph{build efficient machine learning classifiers} based on such training
  examples (Section~\ref{s:prediction}).  The learned data serves as
  guidance for the Monte Carlo tree search that explores different
  proof paths (Section~\ref{sec:proofsearch}). The result, if
  successful, is a verified human-level proof composed of \holfour
  tactics (Section~\ref{sec:proofdisplay}). We evalute the system on a large
  set of theorems originating
  from \holfour (Section~\ref{s:experiments}).

@@

We optimize \tactictoe by tuning parameters of six different techniques:
the timeout of tactics, orthogonalization, abstraction, MCTS policy, MCTS
evaluation and ATP integration. During training, optional techniques such as
orthogonalization may also be turned off completely.
\tactictoe also includes other important techniques which are run with their
default parameters: feature extraction, prediction
algorithms, number of premises for ATPs and MCTS evaluation radius (set by default to
10).
By choosing a set of parameters, we create a strategy for
\tactictoe that is evaluated on our training set of 273 theorems. And the
strategy with the highest number of re-proven theorems is selected for
a full scale evaluation.
